Curiosity-Driven Development of Tool Use Precursors: a Robotic Model

### 2016-02-01 cogsci2016/paper.tex
typos

	The understanding of tool use development in young children is one of the key question for the more general understanding of the ontogeny of human cognition.
	Indeed, a series of abilities are progressively developed from the simplest reaching movements of the arms through more dexterous manipulation of a spoon, towards advanced control of multiple interacting objects.
	The latter shows an understanding of shapes, forces and other physical properties that can be recruited for mental transformations and planning operations which are pillars of human cognition.
	Children's development has first been described as staircase-like successive stages in which all children go through \cite{piaget1952origins}.
	More recently, different views were developed describing variability among children's developmental paths. 
	Siegler's overlapping waves theory \cite{siegler1996emerging} also describes variability in a child's set of current methods to solve a problem, and documents their evolution. 
	In particular, the development of tool use precursors can be described as three consecutive and overlapping stages of behaviours \cite{guerin2013survey}: 
	body babbling, behaviours with a single object, and behaviours with several interacting objects.
	A study of free play \cite{Zelazo198095} shows that at $9\frac{1}{2}$ months play is mostly composed of tactile examination, waving, banging and mouthing of a single object, 
	but that simple relational acts of banging two objects together is already common.
	Later at $13\frac{1}{2}$ months, the study reveals that most children instead prefer to explore the relationships
	among objects, but still show behaviors of the previous phase. 
	Furthermore, they show that this overlapping phases pattern averaged across children is also present in a longitudinal study of a single child.

	In this paper we focus on the study of this progression between overlapping phases of behaviours with objects in a robotic model and in particular on the use of concurrent methods to solve a problem.	
	We hypothesize that several mechanisms play a role in the structure of this behavioural progression and in particular 
	1) the intrinsic motivation to explore as a self-regulation of the learning growth of complexity, and 
	2) the structure of the representation used to encode sensorimotor experience.	
	Active learning is a paradigm where the learning agent chooses which actions to perform in its environment.
	To make this choice the agent can optimize some extrinsic rewards given by the environment, as in standard Reinfocement Learning tasks.
	These extrinsic rewards are not suited for an open-ended autonomous exploration of the environment that could reflect children learning by free play.
	Based on infant's developmental psychology, different types of intrinsic motivations have been studied \cite{santucci2013}.
	One way to define an intrinsic motivation is to autonomously search for situations where the learning progress is high, so as to avoid unreachable or unlearnable situations.
	Computational models have shown that developmental trajectories could emerge from the curiosity-driven learning of sensorimotor mappings, in very different settings.
	In the Playground Experiment \cite{oudeyer_what_2007}, a quadruped robot learned how to use its motor primitives to interact with the items of an infant play mat and a robot peer.
	Also, in a study of the self-organization of vocalizations \cite{moulin-frier_self-organization_2014}, an agent had to learn how to
	use a vocal synthesizer by self-exploration or with the help of humans' demonstrations of phonetic items. 
	This model reproduces accurately major phases of infant vocal development until $6$ months.
	In both studies, developmental trajectories of increasing complexity are emerging from learning, with both regularities in developmental steps and diversity.
	The diversity comes from different mechanisms: stochasticity in the algorithms, variability in the environment, and the multiples attractors of the dynamic learning system.
	In existing models, the agent learns only one mapping that relates a motor space to a single task space. 
	However, in the perspective of an open-ended development of reusable skills, and specifically in the development of tool use, multiple interdependent and hierarchically organized task spaces should be available to the agent.
	For instance, using a tool to act upon an object could make use of previously explored parameterized interaction with the tool.

	We study aspects of those hypothesis leveraging previous models of curiosity-driven learning and extending them to the active exploration of hierarchical sensorimotor and task spaces.
	We define hierarchies of sensorimotor models that structure the sensory space to reflect the interaction of the different items of the environment.
	In such hierarchies of models to explore, different exploration choices are available to the agent at each learning iteration: which model to explore, and how to explore that model.
	The problem of finding an efficient active choice strategy is an instance of strategic learning \cite{nguyen2012}, 
	where different outcomes and strategies are available and the agent has to learn which strategies are useful for which outcomes. 
	This can be viewed as a generalization of active learning methods in machine learning.
	We define the HACOB (Hierarchical Active Curiosity-driven mOdel Babbling) architecture and compare several possible strategies 
	to study the role of active learning and hierarchical representation in the structuration of developmental trajectories.
	We compare the different learning conditions in a $2$D environment where a simulated arm with three joints plus a gripper can grab one of two available tools to move an out-of-reach object.
	We measure the different phases of behaviour during exploration and compare the different structures of behavioural evolution.

@@

	However, here we do not study some other important factors in the development of tool use.
	We consider the hierarchy of sensorimotor models to learn as given to the agent and do not study its autonomous building and evolution.
	Also, social guidance with imitation and mimicry is of central importance for the development of tool use in infants but we do not address the question
	of its modeling in this paper nor the interplay and tradeoff between social learning and self-exploration.
	Another important feature of young infants' learning of tool use is the need to adapt to a developing body and to the maturation of vision during ontogenetic development, 
	but here we consider motor control and sensory perception steady over the simulated learning time span.

@@

		We simulate a 2D robotic arm that can grasp tools that can be used to move an object into different boxes in the environment. 		
		In each trial, the agent executes a motor trajectory and gets the associated sensory feedback. 
		Finally the arm, tools and objects are resetted to their intial state.
		The next sections precisely describe the items of the environment and their interactions.	
		See Fig.\ref{env} for an exemple state of the environment.

@@

			At the end of the movement, the robot gets sensory feedback from the different items of the environment.
			It gets the trajectory of the gripper ($S_{Hand}$, $9$D), the trajectory of the end of the sticks ($S_{Stick_1}, $6$D$ and $S_{Stick_2}, $6$D$), 
			the position of the object at the end of the mouvement ($S_{Object}, $2$D$), and whether the object is in a box at the end of the mouvement and the distance between the object and the nearest box ($S_{Boxes}, $2$D$).		
			The trajectory of the gripper is represented as the $x$ and $y$ positions and the aperture ($1$ or $-1$) of the gripper at $3$ time points: 
			steps $17$, $33$, $50$ during the movement of $50$ steps ($9$D).
			Similarly, the trajectories of the end points of the sticks are $3$-points sequences of $x$ and $y$ positions ($6$D for each stick).
			The agent receives the identifier (from $1$ to $10$) of the reached box if one of them has been reached by the sphere, 0 otherwise. 
			He also gets the minimal distance of the object (at the end of the movement) to the center of a box, even if none have been reached.			 
			The sensory information thus contains $9$ values for the trajectory of the gripper, $6$ for the trajectory of the end of each stick, $2$ for the end position of the object and $2$ for the boxes.
			The total sensory space $S$ has $25$ dimensions.

@@

	Fig. \ref{res_interests} shows details about one trial of the condition H-P-AMB. 
	We can see the interest of each model during the whole experiment.
	The interests of models 2 and 5 increase abruptly once the arm succeeded in grabbing the corresponding stick.
	Following that, the interests of models 3 and 6 increase abruptly once the object has been touched by the corresponding stick.
	An exemple of exploration of the $2$D space of the object is also provided in Fig. \ref{res_interests}(b) for the same condition.

@@

		We provide a measure of the different types of behaviours with the sticks and the object during exploration. 
		We categorize the behaviours into three types.
		In the first category ($hand$) are mouvements of the arm that did not grab any stick and thus did not move the out-of-reached object.
		The second category ($stick$) are mouvements that did grab one of the two sticks but did not touch the object with it.
		The third category ($object$) contains the mouvements where both a stick was grabbed and the object was moved by the stick.
		Fig. \ref{res_ow} shows the prototypical evolution of the proportion of the three categories of behaviour along the $100000$ iterations for conditions H-GR-AMB and H-P-AMB.
		We performed a more detailed analysis by counting the trials where the evolution of the three types of behaviours were similar to the one of Fig. \ref{res_ow}(b).
		A structure was considered similar if it validated each of the following criteria:
		behaviours of categories $stick$ and $object$ increase quikly from $0$ to more than $10\%$ (potentially after an initial phase with a steady low value), 
		and are followed by a plateau (steady curve with small slope) with no abrupt changes, and behaviours of category $object$ start to raise at least $1000$ iterations after $stick$ started to raise.
		See Table \ref{table} for the results of this analysis per condition.
		Also, the median number of abrupt changes across trials for each condition are reported in the same table (as the sum of steady changes of more than $10\%$ in the three behaviours).
		with a significant difference between condition H-GR-AMB and the others (Mann-Whitney U tests, $p<0.0001$).

@@

		\begin{figure*}[ht]
			\centering
			\includegraphics[width=5.5cm]{./include/F-RGB-log53-events-100000.pdf}
			\includegraphics[width=5.5cm]{./include/H-RGB-GR-AMB-log13-events-100000.pdf}
			\includegraphics[width=5.5cm]{./include/H-RGB-P-AMB-log38-events-100000.pdf}
			\caption{Behavioural phases typical exemples. Left: F-RGB. Middle: H-GR-AMB, Right: H-P-AMB. Only hierarchical active condition H-P-AMB shows overlapping phases of behaviour similar to infant development.}
			\label{res_ow}
		\end{figure*}

@@

		Results show different structures of behaviour evolution in the different conditions.		
		Flat architectures do not use the structure of the environment and do not separate the high-dimensional sensorimotor space into learnable subspaces.
		Therefore, they do not show structure in the behavioural evolution but rather steady proportions of the three behaviours.
		By contrast, hierarchical condition H-GR-AMB shows successive behavioural steps with abrupt changes, which reflects the greedy choice of model to babble.
		When one model becomes more interesting than another, it is chosen
		for a large number of iterations until another model exceeds its interest.
		However, random model babbling shows overlapping phases structures in the evolution of the three behaviours more compatible with infants' studies ($60\%$), 
		but less than active model babbling ($70\%$ and $79\%$).
		This is because random model babbling does not adapt its choice of models to their interests along development.
		Indeed, it often explores model $1$ even if it is sufficiently explored to make progress on higher-level models, 
		and so explores less the object position space than active model babbling (H-P-AMB).
		Also, in this setup, all models are still useful to explore after the number of iterations simulated 
		here so the first behavioural phases ($hand$ and $stick$) do not lessen towards the end of the simulations in active conditions. 
		Consequently, the difference between active and random model babbling is more on the cognitive or intentional level, 
		as active model babbling monitors the current progress of each model whereas random model babbling do not, than on a quantitative level.
		However, in others setups where some tasks are learned much faster than others and where at some point it become useless to explore a mastered task, active model babbling would
		make a key difference on a quantitative exploration point of view and on the structure of the evolution of the measured behaviours.\\

@@

		The comparison of conditions H-P-AMB and H-P-AMB-ATC shows that when the agent chooses the tool to reach a given object goal position based on the interest of the corresponding models,
		both tools are trained to reach all goals instead of training only the best performing tool.
		Indeed, we can see that with the active choice of tool, the small stick has produced more diverse effects on the object than in the best choice of tool's condition, even if those effects could also have been 
		generated with the long tool.
		The active tool choice is in accordance with Siegler's overlapping waves theory that documents the parallell use of non-optimal strategies in infants in different contexts
		and explains that a non-optimal strategy gains to be refined as it might turn to be finally a good one for this or for different but related problems.

>>>>

	The understanding of tool use development in young children is one of the key question for the more general understanding of the ontogeny of human cognition.
	Indeed, a series of abilities are progressively developed from the simplest reaching movements of the arms through more dexterous manipulation of a spoon, towards advanced control of multiple interacting objects.
	The latter shows an understanding of shapes, forces and other physical properties that can be recruited for mental transformations and planning operations which are pillars of human cognition.
	Child development has first been described as staircase-like successive stages in which all children go through \cite{piaget1952origins}.
	More recently, different views were developed describing variability among children's developmental paths. 
	Siegler's overlapping waves theory \cite{siegler1996emerging} also describes variability in a child's set of current methods to solve a problem, and documents their evolution. 
	In particular, the development of tool use precursors can be described as three consecutive and overlapping stages of behaviours \cite{guerin2013survey}: 
	body babbling, behaviours with a single object, and behaviours with several interacting objects.
	A study of free play \cite{Zelazo198095} shows that at $9\frac{1}{2}$ months play is mostly composed of tactile examination, waving, banging and mouthing of a single object, 
	but that simple relational acts of banging two objects together is already common.
	Later at $13\frac{1}{2}$ months, the study reveals that most children instead prefer to explore the relationships
	among objects, but still show behaviors of the previous phase. 
	Furthermore, they show that this overlapping phases pattern averaged across children is also present in a longitudinal study of a single child.

	In this paper we focus on the study of this progression between overlapping phases of behaviours with objects in a robotic model and in particular on the use of concurrent methods to solve a problem.	
	We hypothesize that several mechanisms play a role in the structure of this behavioural progression and in particular 
	1) the intrinsic motivation to explore as a self-regulation of the learning growth of complexity, and 
	2) the structure of the representation used to encode sensorimotor experience.	
	Active learning is a paradigm where the learning agent chooses which actions to perform in its environment.
	To make this choice the agent can optimize some extrinsic rewards given by the environment, as in standard Reinforcement Learning tasks.
	These extrinsic rewards are not suited for an open-ended autonomous exploration of the environment that could reflect children learning by free play.
	Based on infant's developmental psychology, different types of intrinsic motivations have been studied \cite{santucci2013}.
	One way to define an intrinsic motivation is to autonomously search for situations where the learning progress is high, so as to avoid unreachable or unlearnable situations.
	Computational models have shown that developmental trajectories could emerge from the curiosity-driven learning of sensorimotor mappings, in very different settings.
	In the Playground Experiment \cite{oudeyer_what_2007}, a quadruped robot learned how to use its motor primitives to interact with the items of an infant play mat and a robot peer.
	Also, in a study of the self-organization of vocalizations \cite{moulin-frier_self-organization_2014}, an agent had to learn how to
	use a vocal synthesizer by self-exploration or with the help of humans' demonstrations of phonetic items. 
	This model reproduces accurately major phases of infant vocal development until $6$ months.
	In both studies, developmental trajectories of increasing complexity are emerging from learning, with both regularities in developmental steps and diversity.
	The diversity comes from different mechanisms: stochasticity in the algorithms, variability in the environment, and the multiples attractors of the dynamic learning system.
	In existing models, the agent learns only one mapping that relates a motor space to a single task space. 
	However, in the perspective of an open-ended development of reusable skills, and specifically in the development of tool use, multiple interdependent and hierarchically organized task spaces should be available to the agent.
	For instance, using a tool to act upon an object could make use of previously explored parameterized interaction with the tool.

	We study aspects of those hypothesis leveraging previous models of curiosity-driven learning and extending them to the active exploration of hierarchical sensorimotor and task spaces.
	We define hierarchies of sensorimotor models that structure the sensory space to reflect the interaction of the different items of the environment.
	In such hierarchies of models to explore, different exploration choices are available to the agent at each learning iteration: which model to explore, and how to explore that model.
	The problem of finding an efficient active choice strategy is an instance of strategic learning \cite{nguyen2012}, 
	where different outcomes and strategies are available and the agent has to learn which strategies are useful for which outcomes. 
	This can be viewed as a generalization of active learning methods in machine learning.
	We define the HACOB (Hierarchical Active Curiosity-driven mOdel Babbling) architecture and compare several possible strategies 
	to study the role of active learning and hierarchical representation in the structuring of developmental trajectories.
	We compare the different learning conditions in a $2$D environment where a simulated arm with three joints plus a gripper can grab one of two available tools to move an out-of-reach object.
	We measure the different phases of behaviour during exploration and compare the different structures of behavioural evolution.

@@

	However, here we do not study some other important factors in the development of tool use.
	We consider the hierarchy of sensorimotor models to learn as given to the agent and do not study its autonomous building and evolution.
	Also, social guidance with imitation and mimicry is of central importance for the development of tool use in infants but we do not address the question
	of its modeling in this paper nor the interplay and trade-off between social learning and self-exploration.
	Another important feature of young infants' learning of tool use is the need to adapt to a developing body and to the maturation of vision during ontogenetic development, 
	but here we consider motor control and sensory perception steady over the simulated learning time span.

@@

		We simulate a 2D robotic arm that can grasp tools that can be used to move an object into different boxes in the environment. 		
		In each trial, the agent executes a motor trajectory and gets the associated sensory feedback. 
		Finally the arm, tools and objects are resetted to their initial state.
		The next sections precisely describe the items of the environment and their interactions.	
		See Fig.\ref{env} for an example state of the environment.

@@

			At the end of the movement, the robot gets sensory feedback from the different items of the environment.
			It gets the trajectory of the gripper ($S_{Hand}$, $9$D), the trajectory of the end of the sticks ($S_{Stick_1}, $6$D$ and $S_{Stick_2}, $6$D$), 
			the position of the object at the end of the movement ($S_{Object}, $2$D$), and whether the object is in a box at the end of the movement and the distance between the object and the nearest box ($S_{Boxes}, $2$D$).		
			The trajectory of the gripper is represented as the $x$ and $y$ positions and the aperture ($1$ or $-1$) of the gripper at $3$ time points: 
			steps $17$, $33$, $50$ during the movement of $50$ steps ($9$D).
			Similarly, the trajectories of the end points of the sticks are $3$-points sequences of $x$ and $y$ positions ($6$D for each stick).
			The agent receives the identifier (from $1$ to $10$) of the reached box if one of them has been reached by the sphere, 0 otherwise. 
			He also gets the minimal distance of the object (at the end of the movement) to the center of a box, even if none have been reached.			 
			The sensory information thus contains $9$ values for the trajectory of the gripper, $6$ for the trajectory of the end of each stick, $2$ for the end position of the object and $2$ for the boxes.
			The total sensory space $S$ has $25$ dimensions.

@@

	Fig. \ref{res_interests} shows details about one trial of the condition H-P-AMB. 
	We can see the interest of each model during the whole experiment.
	The interests of models 2 and 5 increase abruptly once the arm succeeded in grabbing the corresponding stick.
	Following that, the interests of models 3 and 6 increase abruptly once the object has been touched by the corresponding stick.
	An example of exploration of the $2$D space of the object is also provided in Fig. \ref{res_interests}(b) for the same condition.

@@

		We provide a measure of the different types of behaviours with the sticks and the object during exploration. 
		We categorize the behaviours into three types.
		In the first category ($hand$) are movements of the arm that did not grab any stick and thus did not move the out-of-reached object.
		The second category ($stick$) are movements that did grab one of the two sticks but did not touch the object with it.
		The third category ($object$) contains the movements where both a stick was grabbed and the object was moved by the stick.
		Fig. \ref{res_ow} shows the prototypical evolution of the proportion of the three categories of behaviour along the $100000$ iterations for conditions H-GR-AMB and H-P-AMB.
		We performed a more detailed analysis by counting the trials where the evolution of the three types of behaviours were similar to the one of Fig. \ref{res_ow}(b).
		A structure was considered similar if it validated each of the following criteria:
		behaviours of categories $stick$ and $object$ increase quickly from $0$ to more than $10\%$ (potentially after an initial phase with a steady low value), 
		and are followed by a plateau (steady curve with small slope) with no abrupt changes, and behaviours of category $object$ start to raise at least $1000$ iterations after $stick$ started to raise.
		See Table \ref{table} for the results of this analysis per condition.
		Also, the median number of abrupt changes across trials for each condition are reported in the same table (as the sum of steady changes of more than $10\%$ in the three behaviours).
		with a significant difference between condition H-GR-AMB and the others (Mann-Whitney U tests, $p<0.0001$).

@@

		\begin{figure*}[ht]
			\centering
			\includegraphics[width=5.5cm]{./include/F-RGB-log53-events-100000.pdf}
			\includegraphics[width=5.5cm]{./include/H-RGB-GR-AMB-log13-events-100000.pdf}
			\includegraphics[width=5.5cm]{./include/H-RGB-P-AMB-log38-events-100000.pdf}
			\caption{Behavioural phases typical examples. Left: F-RGB. Middle: H-GR-AMB, Right: H-P-AMB. Only hierarchical active condition H-P-AMB shows overlapping phases of behaviour similar to infant development.}
			\label{res_ow}
		\end{figure*}

@@

		Results show different structures of behaviour evolution in the different conditions.		
		Flat architectures do not use the structure of the environment and do not separate the high-dimensional sensorimotor space into learnable sub-spaces.
		Therefore, they do not show structure in the behavioural evolution but rather steady proportions of the three behaviours.
		By contrast, hierarchical condition H-GR-AMB shows successive behavioural steps with abrupt changes, which reflects the greedy choice of model to babble.
		When one model becomes more interesting than another, it is chosen
		for a large number of iterations until another model exceeds its interest.
		However, random model babbling shows overlapping phases structures in the evolution of the three behaviours more compatible with infants' studies ($60\%$), 
		but less than active model babbling ($70\%$ and $79\%$).
		This is because random model babbling does not adapt its choice of models to their interests along development.
		Indeed, it often explores model $1$ even if it is sufficiently explored to make progress on higher-level models, 
		and so explores less the object position space than active model babbling (H-P-AMB).
		Also, in this setup, all models are still useful to explore after the number of iterations simulated 
		here so the first behavioural phases ($hand$ and $stick$) do not lessen towards the end of the simulations in active conditions. 
		Consequently, the difference between active and random model babbling is more on the cognitive or intentional level, 
		as active model babbling monitors the current progress of each model whereas random model babbling do not, than on a quantitative level.
		However, in others setups where some tasks are learned much faster than others and where at some point it become useless to explore a mastered task, active model babbling would
		make a key difference on a quantitative exploration point of view and on the structure of the evolution of the measured behaviours.\\

@@

		The comparison of conditions H-P-AMB and H-P-AMB-ATC shows that when the agent chooses the tool to reach a given object goal position based on the interest of the corresponding models,
		both tools are trained to reach all goals instead of training only the best performing tool.
		Indeed, we can see that with the active choice of tool, the small stick has produced more diverse effects on the object than in the best choice of tool's condition, even if those effects could also have been 
		generated with the long tool.
		The active tool choice is in accordance with Siegler's overlapping waves theory that documents the parallel use of non-optimal strategies in infants in different contexts
		and explains that a non-optimal strategy gains to be refined as it might turn to be finally a good one for this or for different but related problems.
